### 2016-02-25 Appendix.tex

\section{Appendix}

\subsection{System of Units and Metric}
In this thesis the natural units are used, i.e. $c= \hbar (= k_B) = 1$. Furthermore the Minkowski metric is chosen to be
\begin{align}
g^{\mu\nu} = \mathrm{diag}(1, -1, -1, -1).
\end{align}

\subsection{Colour Algebra $SU(N)$}
[Marina von Steinkirch]\\
Tha Casimir operator $C(R)\mathbbm{1}$ of a semi-simple Lie algebra in the irreducible representation $R$ is given by
\begin{align}
g^{ab} T^a(R)T^b(R) = C(R)\mathbbm{1} \label{eq:Casimir}
\end{align}
where $T^a(R)$ is the $a$-th generator of the matrix valued representation $R$, $g^{ab}$ is the metric of group, $C(R)$ is the Quadratic Casimir invariant of the representation $R$ and $\mathbbm{1}$ is the identity in the representation space.\\
Apart from $C(R)$ it is common to define the Dynkin-Index $T(R)$:
\begin{align}
\mathrm{Tr}\left[T^a(R)T^b(R)\right] = T(R)\delta^{ab}.
\end{align}
The two constants are connected by
\begin{align}
C(R) \cdot \mathrm{dim}(R) = T(R) \cdot \mathrm{dim}(G)
\end{align}
where $\mathrm{dim}(G)$ is the dimension of the group and $\mathrm{dim}(R)$ is the dimension of the irreducible representation $R$.\\
In the case of $SU(N)$ one has a diagonal metric $g^{ab} = \delta^{ab}$ and therefore \ref{eq:Casimir} turns to
\begin{align}
\sum_a (T^a(R))^2 = C(R)\mathbbm{1}_{\mathrm{dim}(R) \times \mathrm{dim}(R)}
\end{align}
and one can write down the following useful formulae for the fundamental representation $R=F$: $T^a_{ij} = \frac{\lambda^a_{ij}}{2}$ and the adjoint representation $R=A$: $(T^a_{ij})^{adj} = -if_{aij}$
\begin{align}
T^a_{ik} T^a_{kj} &= C(F)\mathbbm{1}_{ij} && \mbox{with}\ C(F) = \frac{N^2-1}{2N} = \frac{4}{3}\nonumber\\
f^{abc} f^{dbc} &= C(A)\delta^{ad} && \mbox{with}\ C(A) = N = 3\nonumber\\
\mathrm{Tr}\left[T^aT^b\right] &= T(F)\delta^{ab} && \mbox{with}\ T(F) = \frac{1}{2}
\end{align}
where $\lambda^a_{ij}$ are for $N_c = 3$ the Gell-Mann matrices and $f_{abc}$ are the structure constants of $SU(N_c)$.
%\begin{tabular}{c|c|c}
%Quadratic Casimir Invariant and $T(R)$ & $SU(N)_c$ & $SU(3)_c$\\
%\hline
%$C_2(3)$ & $\frac{N^2-1}{2N}$ & $\frac{4}{3}$\\
%$C_2(8)$ & $N$ & $3$\\
%$T(3)$ & $\frac{1}{2}$ & $\frac{1}{2}$ \\
%$T(8)$ & $N$ & $3$ 
%\end{tabular}

\subsection{Weyl basis and 2-spinor notation}\label{sec:2spinor_notation}
As representation of the $\gamma$-matrices the Weyl or chiral representation is chosen:
\begin{align}
&\gamma^\mu = \begin{pmatrix}
0 && \sigma^\mu\\
\overline{\sigma}^\mu && 0
\end{pmatrix} &&
\gamma_5 = \begin{pmatrix}
-\mathbbm{1}_2 && 0\\
0 && \mathbbm{1}_2
\end{pmatrix}
\end{align}
with
\begin{align}
&\sigma^\mu = \begin{pmatrix}
\mathbbm{1}_2 && \sigma^i
\end{pmatrix}, &&
\overline{\sigma}^\mu = \begin{pmatrix}
\mathbbm{1}_2 && -\sigma^i
\end{pmatrix},\label{eq:Pauli}
\end{align}
where $\sigma^i$ are the Pauli matrices and $\mathbbm{1}_n$ is the $n$x$n$ unit matrix. The left and right handed projectors are then given by
\begin{align}
P_L = \frac{1}{2}(\mathbbm{1}_4 - \gamma_5)\begin{pmatrix}
\mathbbm{1}_2 && 0\\
0 && 0
\end{pmatrix} && 
P_R = \frac{1}{2}(\mathbbm{1}_4 + \gamma_5)\begin{pmatrix}
0 && 0\\
0 && \mathbbm{1}_2
\end{pmatrix}\label{eq:projectors}
\end{align}
The generator of the Lorentz group on 4-spinor space is composed of the above matrices. Because of the block form of those it is not surprising that the representation on 4-spinor space is reducible to two representations on 2-spinor (Weyl spinor) spaces. It is therefore sensible to decompose a 4 spinor into a left and a right handed Weyl spinor\footnote{The projectors in the chiral basis \ref{eq:projectors} explain the names left and right handed Weyl spinors.}
\begin{align}
\Psi = \begin{pmatrix}
\psi_\alpha \\
\overline{\chi}^{\dot{\alpha}}
\end{pmatrix}
\end{align}
where $\alpha,\dot{\alpha} = 1,2$. Left handed Weyl spinors are labeled with undotted and right handed Weyl spinors with dotted indices. One distinguishes 4 different Weyl spinors:
\begin{align}
\psi^\alpha, \hspace{1cm} 
\overline{\psi}^{\dot{\alpha}} = (\psi^\alpha)^\ast, \hspace{1cm} \psi_\alpha = \epsilon_{\alpha\beta}\psi^\beta, \hspace{1cm} \mathrm{and} \hspace{1cm} \psi_{\dot{\alpha}} = \epsilon_{\dot{\alpha}\dot{\beta}}\psi^{\dot{\beta}} = (\psi_\alpha)^\ast,
\end{align}
where $^\ast$ denotes complex conjugation and indices are lowered with the antisymmetric $\epsilon_{\alpha\beta}$ ($\epsilon_{\dot{\alpha}\dot{\beta}}$), which obeys
\begin{align}
\epsilon^{\alpha\beta} = \epsilon_{\beta\alpha},\ \epsilon^{\dot{\alpha}\dot{\beta}} = \epsilon_{\dot{\beta}\dot{\alpha}} \hspace{3cm} \mathrm{and} \hspace{3cm} \epsilon_{12} = \epsilon_{\dot{1}\dot{2}}= 1.
\end{align}
By virtue of the antisymmetry of $\epsilon$ one has for the Lorentz invariant product:
\begin{align}
\psi\chi &:= \psi^\alpha\chi_\alpha = -\chi_\alpha\psi^\alpha = \chi^\alpha\psi_\alpha = \chi\psi,\nonumber\\
\overline{\psi}\overline{\chi} &:= \overline{\psi}_{\dot{\alpha}} \overline{\chi}^{\dot{\alpha}} = -\overline{\chi}^{\dot{\alpha}}\overline{\psi}_{\dot{\alpha}} = \overline{\chi}_{\dot{\alpha}}\overline{\psi}^{\dot{\alpha}} = \overline{\chi}\overline{\psi} \label{eq:Weylspinorproduct}
\end{align}
To make the index structure of the Pauli matrices explicit one writes $\sigma^\mu_{\alpha\dot{\alpha}}$ and $\sigma^{\mu\dot{\alpha}\alpha}$ for the formulae in \ref{eq:Pauli}. For the definition of the super algebra in ??? the generators of the Lorentz group on the left and right handed Weyl spinor space are introduced:
\begin{align}
\frac{1}{2}(\sigma^{\mu\nu})_\alpha^{\ \beta} := \frac{i}{4}(\sigma^\mu\overline{\sigma}^\nu - \sigma^\nu\overline{\sigma}^\mu)_\alpha^{\ \beta}\nonumber\\
\frac{1}{2}(\overline{\sigma}^{\mu\nu})^{\dot{\alpha}}_{\ \dot{\beta}} := \frac{i}{4}(\overline{\sigma}^\mu\sigma^\nu - \overline{\sigma}^\nu\sigma^\mu)^{\dot{\alpha}}_{\ \dot{\beta}}
\end{align}

With the definition of bared and charge conjugated 4-spinors\footnote{$\Psi^T$ denotes the transpose of the spinor $\Psi$ and $\Psi^\dagger$ is the Hermitian adjoint of $\Psi$.}
\begin{align}
&\overline{\Psi} := \Psi^\dagger \gamma^0, && \Psi^C := i\gamma^2\gamma^0 \overline{\Psi}^T  
\end{align}
one obtains:
\begin{align}
\Psi &= \begin{pmatrix}
\psi_\alpha \\
\overline{\chi}^{\dot{\alpha}}
\end{pmatrix}, && \ \ 
\overline{\Psi} = \begin{pmatrix}
\chi^\alpha & \overline{\psi}_{\dot{\alpha}}
\end{pmatrix},\nonumber\\
\Psi^C &= \begin{pmatrix}
\chi_\alpha \\
\overline{\psi}^{\dot{\alpha}}
\end{pmatrix}, && 
\overline{\Psi}^C = \begin{pmatrix}
\psi^\alpha & \overline{\chi}_{\dot{\alpha}}
\end{pmatrix}.
\end{align}
The 4-spinor of an arbitrary quark $q$ is given in terms of Weyl spinors $q_L$ and $\overline{q}_R$ by
\begin{align}
q = \begin{pmatrix}
q_L \\
\overline{q}_R
\end{pmatrix}
\end{align}
whereas the 4-spinor of the Dirac gauginos is given in terms of the Weyl spinors $\lambda$ and $\overline{\chi}$\footnote{$\lambda$ is the superpartner of the gluon, called the gluino and $\overline{\chi}$ is the Weyl spinor of the chiral superfield which is associated with the gluon, called the octino.}
\begin{align}
\tilde{g}^a = \begin{pmatrix}
-i \lambda^a \\
i \overline{\chi}^a
\end{pmatrix}
\end{align}

\subsection{Anticommuting numbers}
Anticommuting numbers $\theta^\alpha$ are also referred to as Grassmann numbers and are defined by $\theta^\alpha\theta^\beta = -\theta^\beta\theta^\alpha$ and commute with ordinary numbers.\\
They occur in superspace formalism in the form of 2 tuples, i.e. $\theta^\alpha$ with $\alpha=1,2$. The complex conjugate of this tuple is denoted with $\overline{\theta}^{\dot{\alpha}}$. 
Derivatives are defined by
\begin{align}
\partial^\alpha \theta_\beta := \frac{\partial}{\partial \theta_\alpha} \theta_\beta := \delta^\alpha_\beta \hspace{4cm} \partial_\alpha \theta^\beta := \frac{\partial}{\partial \theta^\alpha} \theta^\beta := \delta_\alpha^\beta\nonumber\\
\overline{\partial}^{\dot{\alpha}} \overline{\theta}_{\dot{\beta}} := \frac{\partial}{\partial \overline{\theta}_{\dot{\alpha}}} \overline{\theta}_{\dot{\beta}} := \delta^{\dot{\alpha}}_{\dot{\beta}} \hspace{4cm} \overline{\partial}_{\dot{\alpha}} \overline{\theta}^{\dot{\beta}} := \frac{\partial}{\partial \overline{\theta}^{\dot{\alpha}}} \overline{\theta}^{\dot{\beta}} := \delta_{\dot{\alpha}}^{\dot{\beta}}
\end{align}
whereby one needs to be cautious as these definitions imply
\begin{align}
\partial_\alpha = -\epsilon_{\alpha\beta}\partial^\beta \hspace{4cm} \partial_{\dot{\alpha}} = -\epsilon_{\dot{\alpha}\dot{\beta}}\partial^{\dot{\beta}}.
\end{align}
Integrals are defined by:
\begin{align}
\int\mathrm{d}\theta_\alpha\ (a+b \theta^\beta + c\theta^\beta \theta^\gamma) := b\delta_{\alpha}^{\beta} + c (\delta_\alpha^\beta \theta^\gamma - \delta_\alpha^\gamma \theta^\beta) \hspace{1cm} \mathrm{and} \hspace{1cm}
\int\mathrm{d}\theta_\alpha\ (a\overline{\theta}^{\dot{\beta}}) := (a\overline{\theta}^{\dot{\beta}}) \int\mathrm{d}\theta_\alpha\
\end{align}
where the first line mirrows the claim of translation invariance. One furthermore introduces the shortcuts
\begin{align}
\int \mathrm{d}\theta^2 = \int\frac{1}{4}\epsilon_{\alpha\beta} \mathrm{d}\theta^\alpha\mathrm{d}\theta^\beta, \hspace{1cm}
\int \mathrm{d}\overline{\theta}^2 = \int\frac{1}{4}\epsilon_{\dot{\alpha}\dot{\beta}} \mathrm{d}\theta^{\dot{\alpha}}\mathrm{d}\theta^{\dot{\beta}}, \hspace{1cm} \mathrm{and} \hspace{1cm}
\int \mathrm{d}^4\theta := \int \mathrm{d}\theta^2\mathrm{d}\overline{\theta}^2
\end{align}

\subsection{Feynman rules for the RSQCD}
The following Feynman rules are derived from \ref{eq:L_RSQCD}. When compared with the Feynman rules of the supersymmetric QCD the diagrams involving scalar gluons are new. In addition the gluon-quark-squark vertex is different in RSQCD for the gauginos are Dirac fermions.\\
refer to paper cited in Beenakker(Appendix: Feynmanrules)
\begin{figure}[!htbp]
\begin{center}
\begin{tikzpicture}[line width=1.5 pt, scale=1.3]
	\node at (-0.5,0.7) {\circled{1}};	
	\draw[scalarnoarrow](0,0)--(1.5,0);
	\node at (180:0.4) {$(\sigma^0)^a$};
	\node at (0:1.9) {$(\sigma^0)^b$};
	\node at (3.4,0) {$\hat{=}\ \frac{i}{p^2-m_{\sigma^0}^2+i\varepsilon}\delta_{ab}$};
\begin{scope}[shift={(7,0)}]
	\node at (-0.5,0.7) {\circled{2}};
	\draw[scalarnoarrow](0,0)--(1.5,0);
	\node at (180:0.4) {$(\phi^0)^a$};
	\node at (0:1.9) {$(\phi^0)^b$};
	\node at (3.4,0) {$\hat{=}\ \frac{i}{p^2-m_{\phi^0}^2+i\varepsilon}\delta_{ab}$};
\end{scope}
\begin{scope}[shift={(0,-2)}]
	\node at (-0.5,0.7) {\circled{3}};
	\draw[scalarbar](0,0)--(1.5,0);
	\node at (180:0.3) {$\tilde{q}_{Ai}$};
	\node at (0:1.8) {$\tilde{q}_{Bj}^\dagger$};
	\node at (3,0) {$\hat{=}\ \frac{i\delta_{AB}}{p^2-m_{\tilde{q}}^2+i\varepsilon}\delta_{ij}$};
\end{scope}
\begin{scope}[shift={(7,-2)}]
	\node at (-0.5,0.7) {\circled{4}};
	\draw[fermionbar](0,0)--(1.5,0);
	\node at (180:0.2) {$q_i$};
	\node at (0:1.7) {$\overline{q}_j$};
	\node at (3,0) {$\hat{=}\ i\frac{\slashed{p}+m_q}{p^2-m_q^2+i\varepsilon}\delta_{ij}$};
\end{scope}
\begin{scope}[shift={(0,-4)}]
	\draw[gluon](1.5,0)--(0,0);
	\node at (-0.5,0.7) {\circled{5}};
	\node at (180:0.2) {$G_\mu^a$};
	\node at (0:1.7) {$G_\nu^b$};
	\node at (3,0) {$\hat{=}\ -i\frac{g^{\mu\nu}}{p^2+i\varepsilon}\delta_{ab}$};
\end{scope}
\begin{scope}[shift={(7,-4)}]
	\node at (-0.5,0.7) {\circled{6}};
	\draw[gluon](1.5,0)--(0,0);
	\draw[fermionnoarrow](0,0)--(1.5,0);
	\node at (180:0.2) {$\tilde{g}^a$};
	\node at (0:1.7) {$\overline{\tilde{g}}^b$};
	\node at (3,0) {$\hat{=}\ i\frac{\slashed{p}+m_{\tilde{g}}}{p^2-m_{\tilde{g}}^2+i\varepsilon}\delta_{ab}$};
\end{scope}
\begin{scope}[shift={(0,-6)}]
	\node at (-0.5,0.7) {\circled{7}};
	\draw[ghostbar](0,0)--(1.5,0);
	\node at (180:0.3) {$c^{a}$};
	\node at (0:1.8) {$\overline{c}^{b}$};
	\node at (3,0) {$\hat{=}\ \frac{i}{p^2+i\varepsilon}\delta_{ab}$};
\end{scope}

\begin{scope}[shift={(1,-8)}]
	\node at (-1.5,0.7) {\circled{8}};
	\draw[fermion] (0,0)--(45:1);
	\draw[fermionbar] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(180:1);
	\node at (45:1.3) {$\overline{q}i$};
	\node at (-45:1.3) {$q_j$};
	\node at (180:1.3) {$G_\mu^a$};
	\node at (2,0) {$\hat{=} -ig_s T^a_{ij}\gamma^\mu$};
\end{scope}
\begin{scope}[shift={(8,-8)}]
	\node at (-1.5,0.7) {\circled{9}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(180:1);
	\node at (45:1.4) {$\tilde{q}^\dagger_{Ai}(-p_1)$};
	\node at (-45:1.4) {$\tilde{q}_{Bj}(p_2)$};
	\node at (180:1.3) {$G_\mu^a$};
	\node at (3,0) {$\hat{=} -ig_s (p_2+p_1)^\mu T^a_{ij}\delta_{AB}$};
\end{scope}
\begin{scope}[shift={(1,-10.5)}]
	\node at (-1.5,0.7) {\circled{10}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(135:1);
	\draw[gluon] (0,0)--(-135:1);
	\node at (45:1.4) {$\tilde{q}^\dagger_{Ai}$};
	\node at (-45:1.4) {$\tilde{q}_{Bj}$};
	\node at (135:1.4) {$G_\mu^a$};
	\node at (-135:1.4) {$G_\nu^b$};
	\node at (3,0) {$\hat{=} \ ig_s^2 g^{\mu\nu} \left\{T^a,T^b\right\}_{ij}\delta_{AB}$};
\end{scope}
\begin{scope}[shift={(1,-13.5)}]
	\node at (-1.5,0.7) {\circled{11}};
	\draw[gluon] (0,0)--(45:1);
	\draw[gluon] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(180:1);
	\node at (45:1.4) {$G_\mu^a(p_a)$};
	\node at (-45:1.4) {$G_\nu^b(p_b)$};
	\node at (180:1.5) {$G_\rho^c(p_c)$};
	\node at (5.5,0) {$\hat{=}  - g_s f_{abc} \left[ g_{\mu\nu}(p_a-p_b)^\rho +\  g_{\nu\rho}(p_b-p_c)^\mu +  g_{\rho\mu}(p_c-p_a)^\nu \right]$};
\end{scope}
\end{tikzpicture}
\caption{In the Feynman diagrams of the propagators the momentum is flowing from the right to the left hand side.\newline
In the Feynman diagrams of the vertices all momenta flow towards the vertex.\newline
The indices $A,B = L,R$ label the right/left-"handedness" of the squarks. The indices $i,j=1,2,3$ are the color indices in the (anti)fundamental representation where $a,b,c,\hdots = 1,\hdots,8$ are the color indices of the adjoint representation. }
\end{center}
\end{figure}
\newpage

%The gluino-quark-squark vertex is not much different from the one in SQCD. One only must avoid replacing $\tilde{g}$ with $\tilde{g}^C$. %See for the derivation of Feynman rules for theories with fermion number violating interaction reference [a. denner, h.eck, o. hahn, j. küblbeck]
\begin{figure}[!htbp]
\begin{center}
\begin{tikzpicture}[line width=1.5 pt, scale=1.3]
	\node at (-1.5,0.7) {\circled{12}};
	\draw[gluon] (0,0)--(45:1);
	\draw[gluon] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(135:1);
	\draw[gluon] (0,0)--(-135:1);
	\node at (45:1.4) {$G_\rho^c$};
	\node at (-45:1.4) {$G_\sigma^d$};
	\node at (135:1.4) {$G_\mu^a$};
	\node at (-135:1.4) {$G_\nu^b$};
	\node at (5.5,0) {$\hat{=} \ -ig_s^2 [f^{abe}f^{cde}(g^{\mu\rho} g^{\nu\sigma} - g^{\mu\sigma}g^{\nu\rho}) + f^{ace}f^{bde}(g^{\mu\nu} g^{\rho\sigma} - g^{\mu\sigma}g^{\nu\rho}) $};
	\node at (8,-0.5) {$+ f^{ade}f^{bce}(g^{\mu\nu} g^{\rho\sigma} - g^{\mu\rho}g^{\nu\sigma})]$};
\begin{scope}[shift={(0,-2.5)}]
	\node at (-1.5,0.7) {\circled{13}};
	\draw[gluon] (0,0)--(180:1);
	\draw[gluon] (0,0)--(45:1);
	\draw[gluon] (0,0)--(-45:1);
	\draw[fermionnoarrow] (0,0)--(-45:1);
	\draw[fermionnoarrow] (0,0)--(45:1);
	\node at (45:1.4) {$\overline{\tilde{g}}^b$};
	\node at (-45:1.4) {$\tilde{g}^c$};
	\node at (180:1.5) {$G_\mu^a$};
	\node(gbar) at (0.9,0.6) {};
	\node(g) at (0.9,-0.6) {};
	\path[line width=0.8pt,->] (g) edge [out=135, in=-135] (gbar);
	\node at (2,0) {$\hat{=}  - g_s f_{abc} \gamma^\mu$};
\end{scope}
\begin{scope}[shift={(7,-2.5)}]
	\node at (-1.5,0.7) {\circled{14}};
	\draw[gluon] (0,0)--(180:1);
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\node at (40:1.4) {$\sigma^{b\dagger}(-p_1)$};
	\node at (-40:1.4) {$\sigma^c(p_2)$};
	\node at (180:1.5) {$G_\mu^a$};
	\node at (3,0) {$\hat{=}  - g_s (p_1+p_2)^\mu f_{abc} $};
\end{scope}
\begin{scope}[shift={(0,-5)}]
	\node at (-1.5,0.7) {\circled{15}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(135:1);
	\draw[gluon] (0,0)--(-135:1);
	\node at (45:1.4) {$\sigma^{\dagger c}$};
	\node at (-45:1.4) {$\sigma^d$};
	\node at (135:1.4) {$G_\mu^a$};
	\node at (-135:1.4) {$G_\nu^b$};
	\node at (3,0) {$\hat{=} \ +ig_s^2 g^{\mu\nu}[f^{aec}f^{bed}+ f^{bec}f^{aed}]$};
\end{scope}
\begin{scope}[shift={(0,-7.5)}]
	\node at (-1.5,0.7) {\circled{16a}};
	\draw[scalar] (0,0)--(45:1);
	\draw[fermionbar] (0,0)--(-45:1);
	\draw[fermionnoarrow] (180:1)--(0,0);
	\draw[gluon] (180:1)--(0,0);
	\node at (45:1.3) {$\tilde{q}_{Li}^\dagger$};
	\node at (-45:1.3) {$\overline{q}^C_j$};
	\node at (180:1.3) {$\tilde{g}^a$};
	\node(g) at (-0.7,-0.3) {};
	\node(u) at (0.4,-0.8) {};
	\path[line width=0.8pt,->] (g) edge [out=0, in=135] (u);
	\node at (3,0) {$\hat{=} -i\sqrt{2}g_s T^a_{ij}P_L$};
\end{scope}
\begin{scope}[shift={(7,-7.5)}]
	\node at (-1.5,0.7) {\circled{16b}};
	\draw[fermion] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[fermionnoarrow] (180:1)--(0,0);
	\draw[gluon] (180:1)--(0,0);
	\node at (45:1.3) {$q^C_i$};
	\node at (-45:1.3) {$\tilde{q}_{Lj}$};
	\node at (180:1.3) {$\overline{\tilde{g}}^a$};
	\node(g) at (-0.7,0.3) {};
	\node(u) at (0.4,0.8) {};
	\path[line width=0.8pt,->] (g) edge [out=0, in=225] (u);
	\node at (3,0) {$\hat{=} -i\sqrt{2}g_s T^a_{ij}P_R$};
\end{scope}
\begin{scope}[shift={(0,-10)}]
	\node at (-1.5,0.7) {\circled{17a}};
	\draw[scalarbar] (0,0)--(45:1);
	\draw[fermion] (0,0)--(-45:1);
	\draw[fermionnoarrow] (180:1)--(0,0);
	\draw[gluon] (180:1)--(0,0);
	\node at (45:1.3) {$\tilde{q}_{Rj}$};
	\node at (-45:1.3) {$\overline{q}_i$};
	\node at (180:1.3) {$\tilde{g}^a$};
	\node(g) at (-0.7,-0.3) {};
	\node(u) at (0.4,-0.8) {};
	\path[line width=0.8pt,->] (g) edge [out=0, in=135] (u);
	\node at (3,0) {$\hat{=} +i\sqrt{2}g_s T^a_{ij}P_L$};
\end{scope}
\begin{scope}[shift={(7,-10)}]
	\node at (-1.5,0.7) {\circled{17b}};
	\draw[fermionbar] (0,0)--(45:1);
	\draw[scalar] (0,0)--(-45:1);
	\draw[fermionnoarrow] (180:1)--(0,0);
	\draw[gluon] (180:1)--(0,0);
	\node at (45:1.3) {$q_j$};
	\node at (-45:1.3) {$\tilde{q}_{Ri}^\dagger$};
	\node at (180:1.3) {$\overline{\tilde{g}}^a$};
	\node(g) at (-0.7,0.3) {};
	\node(u) at (0.4,0.8) {};
	\path[line width=0.8pt,->] (g) edge [out=0, in=225] (u);
	\node at (3,0) {$\hat{=} +i\sqrt{2}g_s T^a_{ij}P_R$};
\end{scope}
\begin{scope}[shift={(0,-12.5)}]
	\node at (-1.5,0.7) {\circled{18a}};
	\draw[scalar] (0,0)--(45:1);
	\draw[fermionnoarrow] (0,0)--(-45:1);
	\draw[gluon] (0,0)--(-45:1);
	\draw[fermionnoarrow] (180:1)--(0,0);
	\draw[gluon] (180:1)--(0,0);
	\node at (45:1.3) {$\sigma^{b\dagger}$};
	\node at (-45:1.3) {$\overline{\tilde{g}}^c$};
	\node at (180:1.3) {$\tilde{g}^a$};
	\node(g) at (-0.7,-0.3) {};
	\node(u) at (0.4,-0.8) {};
	\path[line width=0.8pt,->] (g) edge [out=0, in=135] (u);
	\node at (3,0) {$\hat{=} -\sqrt{2}g_s f^{abc}P_L$};
\end{scope}
\begin{scope}[shift={(7,-12.5)}]
	\node at (-1.5,0.7) {\circled{18b}};
	\draw[fermionnoarrow] (0,0)--(45:1);
	\draw[gluon] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[fermionnoarrow] (180:1)--(0,0);
	\draw[gluon] (180:1)--(0,0);
	\node at (45:1.3) {$\tilde{g}^b$};
	\node at (-45:1.3) {$\sigma^c$};
	\node at (180:1.3) {$\overline{\tilde{g}}^a$};
	\node(g) at (-0.7,0.3) {};
	\node(u) at (0.4,0.8) {};
	\path[line width=0.8pt,->] (g) edge [out=0, in=225] (u);
	\node at (3,0) {$\hat{=} +\sqrt{2}g_s f^{abc} P_R$};
\end{scope}
\begin{scope}[shift={(0,-15)}]
	\node at (-1.5,0.7) {\circled{19}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[scalar] (0,0)--(135:1);
	\draw[scalarbar] (0,0)--(-135:1);
	\node at (45:1.4) {$\tilde{q}^\dagger_{Dl}$};
	\node at (-45:1.4) {$\tilde{q}_{Bj}$};
	\node at (135:1.4) {$\tilde{q}^\dagger_{Ck}$};
	\node at (-135:1.4) {$\tilde{q}_{Ai}$};
	\node at (5,0) {$\hat{=} \ -ig_s^2 [T_{ki}^a T_{lj}^a(\delta_{AL}\delta_{CL} - \delta_{AR}\delta_{CR})(\delta_{BL}\delta_{DL}-\delta_{BR}\delta_{DR})$};
	\node at (5.6,-0.5) {$T_{kj}^a T_{li}^a(\delta_{BL}\delta_{CL} - \delta_{BR}\delta_{CR})(\delta_{AL}\delta_{DL}-\delta_{AR}\delta_{DR})]$};
\end{scope}
\end{tikzpicture}
\end{center}
\caption{The curved arrows indicate the fermion flow. The Feynman rules 16b, 17b and 18b are the complex conjugates of 16a, 17a and 18a respectively. Applying a flipping rule to a vertex one has to reverse the curved arrow, i.e. the fermion flow and replace $\Psi$ with $\bar{\Psi}^C$. In addition one has to add a minus sign for Feynman rule 13.}
\end{figure}
\begin{figure}
\begin{center}
\begin{tikzpicture}[line width=1.5 pt, scale=1.3]
	\node at (-1.5,0.7) {\circled{20}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[scalar] (0,0)--(135:1);
	\draw[scalarbar] (0,0)--(-135:1);
	\node at (45:1.4) {$\sigma^{b\dagger}$};
	\node at (-45:1.4) {$\sigma^c$};
	\node at (135:1.4) {$\tilde{q}^\dagger_{Aj}$};
	\node at (-135:1.4) {$\tilde{q}_{Ai}$};
	\node at (3,0) {$\hat{=} \ -g_s^2 T_{ij}^a f^{abc}(\delta_{AL}\delta_{CL} - \delta_{AR}\delta_{BR})$};
\begin{scope}[shift={(0,-2.5)}]
	\node at (-1.5,0.7) {\circled{21}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[scalar] (180:1)--(0,0);
	\node at (45:1.3) {$\tilde{q}^\dagger_{Ai}$};
	\node at (-45:1.3) {$\tilde{q}_{Bj}$};
	\node at (180:1.6) {$\sigma^a+\sigma^{a\dagger}$};
	\node at (3,0) {$\hat{=} -i\sqrt{2}g_s m_{\tilde{g}} T^a_{ij}(\delta_{AL}\delta_{BL}-\delta_{AR}\delta_{BR})$};
\end{scope}
\begin{scope}[shift={(0,-5)}]
	\node at (-1.5,0.7) {\circled{22}};
	\draw[scalar] (0,0)--(45:1);
	\draw[scalarbar] (0,0)--(-45:1);
	\draw[scalar] (0,0)--(135:1);
	\draw[scalarbar] (0,0)--(-135:1);
	\node at (45:1.4) {$\sigma^{\dagger d}$};
	\node at (-45:1.4) {$\sigma^e$};
	\node at (135:1.4) {$\sigma^{\dagger b}$};
	\node at (-135:1.4) {$\sigma^c$};
	\node at (3,0) {$\hat{=} \ -g_s^2 (f^{abc}f^{ade} + f^{abc}f^{adc})$};
\end{scope}
\end{tikzpicture}
\end{center}
\end{figure}

\subsection{Passarino-Veltman Integrals}

\begin{align}
A_0(m^2) &= \frac{i}{16\pi^2} \int_l \frac{1}{l^2-m^2}\nonumber\\
B_{0,\mu,\mu\nu}(p^2,m_1^2,m_2^2) &= \frac{i}{16\pi^2} \int_l \frac{\left\{1,l_\mu,l_\mu l_\nu \right\}}{[l^2-m_1^2][(l+p)^2-m_2^2]}\nonumber\\
C_{0,\mu,\mu\nu}(p_1^2,p_2^2,m_1^2,m_2^2,m_3^2) &= \frac{i}{16\pi^2} \int_l \frac{\left\{1,l_\mu,l_\mu l_\nu \right\}}{[l^2-m_1^2][(l+p_1)^2-m_2^2][(l+p_2)^2-m_3^2]}
\end{align}
with the shortcut $\int_l = \mu^{2\epsilon}\int\frac{\mathrm{d}^D l}{(2\pi)^D}$. Furthermore there are supressed $\varepsilon$'s which prescripe how the poles in the complex plane are avoided. They are hidden in the infinitesimal shift of the masses: $m_i^2 \to m_i^2 - i \varepsilon$.\\
The tensor integrals can be decomposed as
\begin{align}
B_\mu &= p_\mu B_1\nonumber\\
B_{\mu\nu} &= g_{\mu\nu}B_{00} + p_\mu p_\nu B_{11}\nonumber\\
C_\mu &= p_{1\mu}C_1 + p_{2\mu}C_2 \nonumber\\
C_{\mu\nu} &= g_{\mu\nu}C_{00} + p_{1\mu}p_{1\nu}C_{11} + p_{2\mu}p_{2\nu}C_{22} + (p_{1\mu}p_{2\nu} + p_{2\mu}p_{1\nu})C_{12} 
\end{align}
\begin{align}
A_0(m^2) = m^2\left( \Delta -\ln \frac{m^2}{\mu^2} + 1 \right) + \mathcal{O}(\epsilon)
\end{align}
$\Delta = \frac{1}{\epsilon} - \gamma_E+\ln 4\pi$
\begin{align}
B_0(0,m_1^2,m_2^2) &= \frac{A_0(m_1^2)-A_0(m_2^2)}{m_1^2-m_2^2} = \Delta + 1 -\frac{m_1^2\ln \frac{m_1^2}{\mu^2}-m_2^2\frac{m_2^2}{\mu^2}}{m_1^2-m_2^2} + \mathcal{O}(\epsilon)\\
B_0(0,m^2,m^2) &= \frac{\partial}{\partial m^2} A_0(m^2) = \Delta -\ln \frac{m^2}{\mu^2} + 1 + \mathcal{O}(\epsilon)
\end{align}

%\subsection{Cross section and Phase Space Integration}
%\begin{align}
%\mbox{d}\sigma = \frac{1}{2E_A 2E_B |v_A-v_B|}\left( \prod_f \frac{\mbox{d}^3p_f}{(2\pi)^3}\frac{1}{2E_f} \right) \sum |\mathcal{M}(p_A,p_B \to \left\{p_f\right\})|^2 (2\pi)^4 \delta^{(4)}(p_a+p_b-\sum p_f)
%\end{align}
%...\\
%\begin{align}
%\frac{\mbox{d}^2 \sigma}{\mbox{d}t\mbox{d}u} = \frac{K_{ij}}{s^2} \frac{\pi S_{\epsilon}}{\Gamma(1-\epsilon)} \left[ \frac{tu-m_{\tilde{q}}^4}{\mu^2 s}\right]^{-\epsilon} \Theta(tu-m_{\tilde{q}}^4) \Theta(s-4m_{\tilde{q}}^2) \delta(s+t+u-2m_{\tilde{q}}^2) \sum |\mathcal{M}^B|^2
%\end{align}