### 2016-02-25 Masterarbeit.tex

\input{preamble}
\input{custom_commands_Theorie}
\begin{document}

\pagenumbering{gobble}   % no pagenumbering
% Titelpageseite
\begin{titlepage}
 \begin{tabularx}{\linewidth}{X}
  \includegraphics[width=6cm]{TU_Logo_SW} \\\hline\hline
  \vspace{4.5em}
  \begin{singlespace}\begin{center}\bfseries\Huge  
  Title of Master thesis  
  \end{center}\end{singlespace}
  \vspace{5.5em}
  \begin{singlespace}\begin{center}\large
   Master-Arbeit \\ zur Erlangung des Hochschulgrades \\ 
   Master of Science \\ 
   im Master-Studiengang Physik
  \end{center}\end{singlespace}\medskip
  \begin{center}vorgelegt von\end{center}
  \begin{center}
   {\large Sebastian Liebschner} \\ geboren am 08.05.1990 in Meerane
  \end{center}\medskip
  \begin{singlespace}\begin{center}\large
   Institut für Kern- und Teilchenphysik Physik \\
   Fachrichtung Physik \\
   Fakultät Mathematik und Naturwissenschaften \\
   Technische Universität Dresden \\ 2016
  \end{center}\end{singlespace}
 \end{tabularx}
\end{titlepage}

% Gutachterseite
\thispagestyle{empty}\vspace*{48em}
Eingereicht am xx.~Monat~20xx\vspace{1.5em}
\par{\large\begin{tabular}{ll}
 1. Gutachter: & Prof.~Dr.~XX \\
 2. Gutachter: & Prof.~Dr.~YY \\
\end{tabular}}
\newpage

% abstract page
\begin{center}\large\bfseries Summary\end{center}
Abstract \\ 
English: motivation: aesthetic: Coleman-Mandula --> Haag-Lopuszanski-Sohnius-Theorem\\
\vspace{20em}
Abstract \\ 
Deutsch \\
\newpage

\tableofcontents
\newpage

\pagenumbering{arabic}
\section{Introduction}
motivation: aesthetic: Coleman-Mandula --> Haag-Lopuszanski-Sohnius-Theorem\\
plots for exclusion of squarks in specific SUSY scenarios (from Michael) --> R-Symmetry could be possible explanation for that because:\\
MSSM-Lagrangian --> trafo rules for superfields under R-symm --> forbidden terms in MRSSM (write down Lagrangian for R-symmetric SUSYQCD)\\
suppression of squark production in MRSSM by less diagrams ($m_{gluino}^{-4}$ suppression at low energies in MRSSM and only $m_{gluino}^{-2}$ suppression in MSSM)\\
R-charges of all fields (show in diagram!) --> only if R-charges of final / initial particles are zero, a diagram is allowed in R-symm. model
\newpage
\input{Standardmodel}
\newpage
\input{MSSM}
\newpage
\input{RSymmetry}
\newpage
\input{TreeLevel}
\newpage
\input{Renormalization}
\newpage
\input{NLO}
\newpage
\section{Summary and Outlook}
\newpage
\input{Appendix}
\newpage
\bibliography{references}
\newpage
% Erklärung
\clearpage
\thispagestyle{empty}
\minisec{Erklärung}\vspace*{1.5em}
Hiermit erkläre ich, dass ich diese Arbeit im Rahmen der Betreuung am Institut
für ??? Physik ohne unzulässige Hilfe Dritter verfasst und alle Quellen als solche gekennzeichnet habe.
\vspace*{45em}
Vorname Nachname \par
Dresden, Monat 2012

\end{document}