### 2016-04-24 Introduction.tex
Passarino-Veltman integrals and other changes

motivation: aesthetic: Coleman-Mandula --> Haag-Lopuszanski-Sohnius-Theorem\\
plots for exclusion of squarks in specific SUSY scenarios (from Michael) --> R-Symmetry could be possible explanation for that because:\\
MSSM-Lagrangian --> trafo rules for superfields under R-symm --> forbidden terms in MRSSM (write down Lagrangian for R-symmetric SUSYQCD)\\
suppression of squark production in MRSSM by less diagrams ($m_{gluino}^{-4}$ suppression at low energies in MRSSM and only $m_{gluino}^{-2}$ suppression in MSSM)\\
R-charges of all fields (show in diagram!) --> only if R-charges of final / initial particles are zero, a diagram is allowed in R-symm. model\\
references to build in
\begin{itemize}
\item "Matching Squark Pair Production at NLO with Parton Showers" from Gavin, Hangst, Krämer, Mühlleitner,.. for complete treatment of NLO calculation\\
\item "dIRAC gAUGINOS IN susy - sUPPRESSED jETS + <met sIGNALS: a sNOWMASS wHITEPAPER" FROM kRIBS, mARTIN for Squark production at LO, allude to same result\\
\item "Dirac Gaugino Masses and supersoft SUSY breaking" from Fox, Weiner, Nelson for the introduction of the MRSSM
\end{itemize}