### 2016-06-13 MSSM.tex
Latex does not compile any more

\section{The Minimal Supersymmetric Standard Model}
Even though the Standard Model is a very well working machinery which has partially been tested to an impressive accuracy it is not the ultimate theory of nature. This is not only because of its incompleteness regarding gravity but also because of its missing incorporation of dark matter. This is that astrophysical observations \cite{Adam:2015rua} indicate that only 4.9\% of the universe's mass energy consist of ordinary matter while the majority is represented by 26.8\% of dark matter and 68.3\% of dark energy. Another puzzle is the hierarchy problem which questions the large discrepancy between the Higgs boson mass and its quantum correction, see \cite{Martin:1997ns}. Finally the gauge groups of the Standard Model cannot be unified to a theoretically appealing single gauge group at some high energy scale for the three gauge couplings do not meet at any renormalization scale \cite{Martin:1997ns}. There are other problems which are not mentioned here.\\
Supersymmetry provides a possible extension of the Standard Model. It constitute a mathematically very aesthetic solution of these problems. This section explains briefly the mechanism of supersymmetry and the minimal supersymmetric extension of the Standard Model (MSSM).

\subsection{Supersymmetry as an Extention of the Poincaré Symmetry}\label{sec:SUSYalgebra}
As already seen in the previous chapter symmetries are of vital importance when studying particle physics. Coleman and Mandula \cite{Coleman:1967ad} found that for any reasonable quantum field theory spacetime symmetries and internal, i.e. gauge symmetries cannot by combined in any but the trivial way. Ergo the symmetry groups of the Standard Model are the most general ones.\\
This statement was corrected by the Haag Lopuszanski-Sohnius theorem \cite{Haag:1974qh} which weakened the condition of the quantum field theory by allowing also for anticommuting symmetry generators. This is by introducing the generators $Q_\alpha$ and $Q_{\dot{\alpha}}$ with $\alpha, \dot{\alpha} \in \left\{ 1,2 \right\}$ the Poincare-algebra eq. \ref{eq:PoincareAlgebra} can be nontrivially extended to the so called superalgebra (see Appendix \ref{sec:2spinor_notation} for the definition of the $\sigma$ matrices):
\begin{align}
\left\{Q_\alpha,Q_\beta\right\} = \left\{\overline{Q}_{\dot{\alpha}},\overline{Q}_{\dot{\beta}}\right\} &= 0,  \nonumber\\
\left\{Q_\alpha,\overline{Q}_{\dot{\alpha}}\right\} &= 2\sigma^\mu_{\alpha\dot{\alpha}} P_\mu, \nonumber\\
[P^\mu,Q_\alpha] = [P^\mu,\overline{Q}_{\dot{\alpha}}] &= 0, \nonumber \\
[Q_\alpha, J^{\mu\nu}] &= \frac{1}{2} (\sigma^{\mu\nu})_\alpha^{\ \beta} Q_\beta.\label{eq:SUSYalgebra}
\end{align}
As $J^{\mu\nu}$ is the generalized angular momentum operator the last line in eq. \ref{eq:SUSYalgebra} implies that $Q_\alpha$ is a spin $\frac{1}{2}$ operator. Applying $Q_\alpha$ to a field will change the spin of it by $\frac{1}{2}$. This means supersymmetry generators convert fermions to bosons and vice versa.\\
The superalgebra in eq. \ref{eq:SUSYalgebra} is actually the smallest extension of the Poincare-algebra as it is possible to introduce further anticommuting symmetry generators. Their maximal number is fixed to $N = 8$. The MSSM deals only with one set of anticommuting symmetry generators and is therefore referred to be a $N = 1$ supersymmetric theory.\\
A representation in the form of differential operators can be given by introducing superspace. Superspace is a manifold obtained by enlarging Minkowski space, whose coordinates are label with $x^\mu$, with four anticommuting numbers: $\theta^\alpha$ and $\overline{\theta}^{\dot{\alpha}}$, where $\alpha,\ \dot{\alpha} = 1,2$ (see Appendix \ref{sec:AnticommNumbers} for the definition of anticommuting numbers). A representation of the superalgebra reads
\begin{align}
P^\mu &= i\partial^\mu\nonumber\\
J^{\mu\nu} &= i(x^\mu\partial^\nu - x^\nu\partial^\mu)\nonumber\\
Q_\alpha &= i(\partial_\alpha + i\sigma^\mu_{\alpha\dot{\alpha}}
\overline{\theta}^{\dot{\alpha}}\partial_\mu)\nonumber\\
\overline{Q}_{\dot{\alpha}} &= i(-\overline{\partial}_{\dot{\alpha}} - i \theta^\alpha \sigma^\mu_{\alpha\dot{\alpha}}\partial_\mu).\label{eq:SUSYGen}
\end{align}

\subsection{A Generic Supersymmetric Model in Superspace Formulation}
This chapter outlines the generic ingredients and terms of a supersymmetric model. To this end it is practical to work in the language of superspace and superfields. Superspace has just been introduced and superfields are functions on superspace.\\
The for the MSSM relevant superfields\footnote{Superfield are throughout this thesis labeled with a hat.} are the chiral superfield $\hat{\Phi}$, the antichiral superfield $\hat{\overline{\Phi}}$ and the vector superfield $V$. Chiral superfields are defined by the restriction $\overline{\mathcal{D}}_{\dot{\alpha}}\hat{\Phi} = 0$, antichiral superfields by $\mathcal{D}_\alpha\hat{\overline{\Phi}} = 0$ (see Appendix \ref{sec:AnticommNumbers} for the definition of chiral covariant derivatives) and vector superfields by the condition of being real $V^\dagger = V$.\\
Their component decomposition reads\footnote{For tho vector superfield Wess-Zumino-gauge is applied.}
\begin{align}
\hat{\Phi}(x,\theta,\overline{\theta}) &= A(x) + \sqrt{2}\theta\psi(x) + \theta\theta F(x) - i\theta\sigma^\mu \overline{\theta}\partial_\mu A(x) - \frac{1}{4}\theta\theta\overline{\theta}\overline{\theta}\partial_\mu\partial^\mu A(x) - \frac{i}{\sqrt{2}}\theta\theta\overline{\theta}\overline{\sigma}^\mu \partial_\mu\psi(x)\nonumber\\
\hat{\overline{\Phi}}(x,\theta,\overline{\theta}) &= A^\dagger(x) + \sqrt{2}\overline{\theta}\overline{\psi}(x) + \overline{\theta}\overline{\theta} F^\dagger(x) + i\theta\sigma^\mu \overline{\theta}\partial_\mu A^\dagger(x) - \frac{1}{4}\theta\theta\overline{\theta}\overline{\theta}\partial_\mu\partial^\mu A^\dagger(x) - \frac{i}{\sqrt{2}}\overline{\theta}\overline{\theta}\theta\sigma^\mu \partial_\mu\overline{\psi}(x)\nonumber\\
\hat{V}(x,\theta,\overline{\theta}) &= \theta\sigma^\mu\overline{\theta} v_\mu + i\theta\theta\overline{\theta}\overline{\lambda}(x) -i \overline{\theta}\overline{\theta}\theta\lambda(x) + \frac{1}{2}\theta\theta\overline{\theta}\overline{\theta}D(x),\label{eq:superfielddecomp}
\end{align}
where $A(x)$ and $F(x)$ are complex scalar fields, $\psi(x)$ and $\lambda(x)$ are left handed Weyl spinors and $D(x)$ being a real scalar field.\\
The superfields transform under a generic gauge transformation as
\begin{align}
\hat{\Phi} &\to \mathrm{e}^{-2ig\Lambda}\Phi\nonumber\\
\hat{\overline{\Phi}} &\to \hat{\overline{\Phi}}\mathrm{e}^{2ig\overline{\hat{\Lambda}}}\nonumber\\
\mathrm{e}^{2g\hat{V}} &\to \mathrm{e}^{-2ig\hat{\overline{\Lambda}}}\mathrm{e}^{2g\hat{V}} \mathrm{e}^{2ig\hat{\Lambda}},
\end{align}
where $\hat{\Lambda} = \hat{\Lambda}^aT^a$ and $\hat{V} = \hat{V}^aT^a$. $\hat{\Lambda}^a$ is an arbitrary chiral superfield and the $T^a$ are the generetors of the Lie algebra, which is associated to the gauge group in question. g is the gauge coupling constant of the gauge group.\\
One can therefore construct the important gauge invariant term $\int\mathrm{d}^4\theta\ \hat{\overline{\Phi}}\mathrm{e}^{2g\hat{V}}\hat{\Phi}$. If one introduces the gauge covariant derivative $D_\mu = \partial + ig T^a v^a_\mu$ the component decomposition reads
\begin{align}
\int\mathrm{d}^4\theta\hat{\overline{\Phi}}\mathrm{e}^{2g\hat{V}}\hat{\Phi} &= F^\dagger F + \left(D_\mu A\right)^\dagger \left(D^\mu A\right) + \overline{\psi}\overline{\sigma}^\mu i D_\mu \psi\nonumber \\
&- \sqrt{2}g\left( -i(A^\dagger T^aA)\lambda^a +i\overline{\lambda}^a(AT^aA^\dagger) \right) + g(A^\dagger T^a A) D^a .\label{eq:L_matter}
\end{align}
Therefore this term gives rise to the kinetic terms of the components of the chiral and antichiral superfields $A,\ A^\dagger,\ \psi$ and $\overline{\psi}$, their minimal coupling to the gauge fields $v_\mu^a$ and their superpartners $\lambda^a$ and $\overline{\lambda}^a$ and terms involving the auxiliary fields $F,\ F^\dagger$ and $D$.\\
With the field-strength chiral superfields $\hat{W}_\alpha = -\frac{1}{4}\overline{\mathcal{D}}\overline{\mathcal{D}}(\mathrm{e}^{-2gV}\mathcal{D}_\alpha\mathrm{e}^{2gV})$ one can write down a gauge invariant term yielding the kinetic terms of the gauge fields and their superpartners:
\begin{align}
\int\mathrm{d}^2\theta \frac{1}{16g^2 }\hat{W}^{\alpha a} \hat{W}^a_\alpha + h.c.= \frac{1}{2}D^aD^a -\frac{1}{4}F^a_{\mu\nu}F^{a\mu\nu} + \frac{i}{2}\overline{\lambda}^a\overline{\sigma}^\mu(D_\mu\lambda^a) + \frac{i}{2}\lambda^a\sigma^\mu(D_\mu\overline{\lambda}^a).\label{eq:L_gauge}
\end{align}
A third generic term in a supersymmetric theory arises from the superpotential $W(\hat{\Phi})$ which is a holomorphic function in the chiral superfields:
\begin{align}
\int\mathrm{d}^2\theta\ W(\hat{\Phi}).
\end{align}
A renormalizable superpotential is given by $W(\hat{\Phi}) = c_i\hat{\Phi} + \frac{m_{ij}}{2}\hat{\Phi}_i\hat{\Phi}_j + \frac{g_{ijk}}{3!}\hat{\Phi}_i\hat{\Phi}_j\hat{\Phi}_k$. The component decomposition of the corresponding terms is
\begin{align}
\int\mathrm{d}^2\theta\ \hat{\Phi}_1 &= F_1\nonumber\\
\int\mathrm{d}^2\theta\ \hat{\Phi}_1\hat{\Phi}_2 &= A_1F_2 + F_1A_2-\psi_1\psi_2\nonumber\\
\int\mathrm{d}^2\theta\ \hat{\Phi}_1\hat{\Phi}_2\hat{\Phi}_3 &= F_1A_2A_3 + A_1F_2A_3 + A_1A_2F_3 - A_1\psi_2\psi_3 - \psi_1A_2\psi_3 - \psi_1\psi_2A_3.\label{eq:L_superpot}
\end{align}
The Lagrangian for a supersymmetric theory is therefore given by 
\begin{align}
\mathcal{L}_{SUSY} &= \mathcal{L}_{matter} + \mathcal{L}_{gauge} + \mathcal{L}_{superpot}\nonumber\\
&=\int\mathrm{d}^4\theta\hat{\overline{\Phi}}\mathrm{e}^{2g\hat{V}}\hat{\Phi} + \left(\int\mathrm{d}^2\theta \frac{1}{16g^2 }\hat{W}^{\alpha a} \hat{W}^a_\alpha + h.c. \right) +\int\mathrm{d}^2\theta\ W(\hat{\Phi})
\end{align}
Observing the component decomposition \ref{eq:L_matter},\ref{eq:L_gauge}, \ref{eq:L_superpot} of the 3 parts of this Lagrangian, one observes that the $F$ and $D$ fields have no kinetic term and are therefore auxiliary fields which can be eliminated by their equation of motion $\frac{\partial \mathcal{L}}{\partial\phi} = \partial_\mu\frac{\partial \mathcal{L}}{\partial(\partial_\mu \phi)}$ with $\phi = F, D$. Doing this one obtains
\begin{align}
\mathcal{L}_D &= \frac{1}{2} D^aD^a + g A^\dagger T^aD^a A \hspace{1cm} \Rightarrow \hspace{1cm} D^a = - A^\dagger T^a A \nonumber\\
\mathcal{L}_D &= -\frac{1}{2}\left( A^\dagger T^a A \right)^2
\end{align}
and 
\begin{align}
\mathcal{L}_F &= F_i^\dagger F_i + \left( c_iF_i + m_{ij}F_iA_j + \frac{g_{ijk}}{2}F_iA_jA_k + h.c. \right) \hspace{1cm} \Rightarrow \hspace{1cm} F^\dagger_i = - \frac{\partial W(A)}{\partial A_i}\nonumber\\
\mathcal{L}_F &= -\left| \frac{\partial W(A)}{\partial A_i} \right|^2
\end{align}

\subsection{The Minimal Supersymmetric Standard Model in superspace formulation}
The Lagrangian for the MSSM\footnote{This is the Lagrangian on the classical level, i.e. there are neither gauge fixing nor ghost terms.} reads
\begin{align}
\mathcal{L}_{MSSM} = \int \mathrm{d}^4 \theta & \left[ \hat{\overline{Q}}\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V} + 2g_s\hat{V}_s}\hat{Q} + \hat{\overline{U}}\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V} - 2g_s\hat{V}^T_s}\hat{U} + \hat{\overline{D}}\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V} - 2g_s\hat{V}^T_s}\hat{D} \right.\nonumber\\
 & + \hat{\overline{L}}\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V}}\hat{L} + \hat{\overline{E}}\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V}}\hat{E}\nonumber\\
 & + \left. \hat{\overline{H}}_d\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V}}\hat{H}_d + \hat{\overline{H}}_u\mathrm{e}^{2g^\prime \hat{V}^\prime + 2g\hat{V}}\hat{H}_u \right]\nonumber\\
 + \int \mathrm{d}^2\theta & \left[ \frac{1}{16g^{\prime 2}} \hat{W}^{\prime\alpha}\hat{W}^\prime_\alpha + \frac{1}{16g^{2}} \hat{W}^{a\alpha}\hat{W}^a_\alpha + \frac{1}{16g_s^2} \hat{W}_s^{a\alpha}\hat{W}^a_{s\alpha} \right] + h.c.\nonumber\\
 + \int \mathrm{d}^2\theta&\ W_{MSSM} + h.c.\nonumber\\
 +\  \mathcal{L}_{soft}.\ &\label{eq:L_MSSM}
\end{align}
Apart from the already discussed terms in the first 4 lines of \ref{eq:L_MSSM} there is a superpotential $W_{MSSM}$:
\begin{align}
W_{MSSM} &= y_d \hat{H}_d \hat{Q} \hat{D} + y_u \hat{H}_u \hat{Q} \hat U + y_e \hat{H}_d \hat{L} \hat{E} - \mu \hat{H}_d \hat{H}_u\label{eq:W_MSSM}
\end{align}
and terms which break supersymmetry softly, i.e. terms with coupling constants with positive mass dimension.
\begin{align}
\mathcal{L}_{soft} &= -M^2_{\tilde{Q}}|\tilde{q}_L|^2 - M^2_{\tilde{U}}|\tilde{u}_R|^2 - M^2_{\tilde{D}}|\tilde{d}_R|^2 \nonumber\\
&\ \ \ - M^2_{\tilde{L}}|\tilde{l}_L|^2 - M^2_{\tilde{E}}|\tilde{e}_R|^2 - M^2_{H_d}|H_d|^2 - M^2_{H_u}|H_u|^2\nonumber\\
&\ \ \ +\frac{1}{2}\left( M_1 \lambda\lambda + M_2 \lambda^a\lambda^a + M_3 \lambda_s^a\lambda_s^a\right) + h.c.\nonumber\\
&\ \ \ -\left( A_d y_d H_d \tilde{q}_L \tilde{d}^\dagger_R + A_u y_u H_u \tilde{q}_L \tilde{u}^\dagger_R + A_e y_e H_d \tilde{l}_L \tilde{e}^\dagger_R -B\mu H_d H_u \right) + h.c.\label{eq:L_soft}
\end{align}
The field content of the MSSM is summarized in \ref{tab:MSSMfieldcontent}
\begin{table}
\begin{center}
\begin{tabular}{c|c|c}\label{tab:MSSMfieldcontent}
Superfield & Components & $SU_\mathrm{C}(3)$ $\times$ $SU_\mathrm{L}(2)$ $\times$ $U_\mathrm{Y}(1)$\\
\hline
$\hat{\Phi}$ & $A$, $\psi$ & \\
$\hat{V}$ & $\lambda$, $v_\mu$ \\
$\hat{Q}$ & $\tilde{q}_L = \begin{pmatrix}
\tilde{u}_L \\
\tilde{d}_L
\end{pmatrix}$, $q_L = \begin{pmatrix}
u_L \\
d_L
\end{pmatrix}$ & (3, 2, $\frac{1}{6}$)\\
$\hat{U}$ & $\tilde{u}_R^\dagger$, $u_R$ & (3$^\ast$, 1, $-\frac{2}{3}$)\\
$\hat{D}$ & $\tilde{d}_R^\dagger$, $d_R$ & (3$^\ast$, 1, $+\frac{1}{3}$)\\
$\hat{L}$ & $\tilde{l}_L = \begin{pmatrix}
\tilde{\nu}_L \\
\tilde{e}_L
\end{pmatrix}$, $l_L = \begin{pmatrix}
\nu_L \\
e_L
\end{pmatrix}$ & (1, 2, $-\frac{1}{2}$)\\
$\hat{E}$ & $\tilde{e}_R^\dagger$, $e_R$ & (1, 1, 1)\\
$\hat{H}_d$ & $H_d$, $\tilde{H}_d$ & (1, 2, $-\frac{1}{2}$)\\
$\hat{H}_u$ & $H_u$, $\tilde{H}_u$ & (1, 2, $+\frac{1}{2}$)\\
$\hat{V}^\prime$ & $\lambda^\prime$, $B_\mu$ & (1, 1, 0)\\
$\hat{V}^a$ & $\lambda^a$, $W_\mu^a$ & (1, 3, 0)\\
$\hat{V}_s^a$ & $\lambda_s^a$, $G_\mu^a$ & (8, 1, 0)\\
\end{tabular}
\caption{The table shows the field content of the MSSM in terms of the superfields and their component decomposition. The first two lines show the decomposition of the generic superfields (cf. \ref{eq:superfielddecomp}).\newline
The third column shows the representation (for $SU_\mathrm{C}(3)$ and $SU_\mathrm{L}(2)$) in which the fields transform and the charges of the fields for $U_\mathrm{Y}(1)$.}\label{tab:MSSMfieldcontent}
\end{center}
\end{table}