### 2016-07-15 Masterarbeit.tex

\input{preamble}
\input{custom_commands_Theorie}
\begin{document}

\pagenumbering{gobble}   % no pagenumbering
% Titelpageseite
\begin{titlepage}
 \begin{tabularx}{\linewidth}{X}
  \includegraphics[width=6cm]{TU_Logo_SW} \\\hline\hline
  \vspace{4.5em}
  \begin{singlespace}\begin{center}\bfseries\Huge  
  Squark Production in the R-Symmetric Supersymmetric Standardmodel  
  \end{center}\end{singlespace}
  \vspace{5.5em}
  \begin{singlespace}\begin{center}\large
   Masterarbeit \\ zur Erlangung des Hochschulgrades \\ 
   Master of Science \\ 
   im Master-Studiengang Physik
  \end{center}\end{singlespace}\medskip
  \begin{center}vorgelegt von\end{center}
  \begin{center}
   {\large Sebastian Liebschner} \\ geboren am 08.05.1990 in Meerane
  \end{center}\medskip
  \begin{singlespace}\begin{center}\large
   Institut für Kern- und Teilchenphysik \\
   Fachrichtung Physik \\
   Fakultät Mathematik und Naturwissenschaften \\
   Technische Universität Dresden \\ 2016
  \end{center}\end{singlespace}
 \end{tabularx}
\end{titlepage}

% Leerseite
\thispagestyle{empty}
\cleardoublepage

% Gutachterseite
\thispagestyle{empty}\vspace*{48em}
Eingereicht am 15.~Juli~2016\vspace{1.5em}
\par{\large\begin{tabular}{ll}
 1. Gutachter: & Prof.~Dr.~Dominik Stöckinger \\
 2. Gutachter: & Prof.~Dr.~Arno Straessner \\
\end{tabular}}
\newpage

% Leerseite
\thispagestyle{empty}
\cleardoublepage

% abstract page
%\begin{center}\large\bfseries Summary\end{center}
\section*{\centering{Kurzdarstellung}}
Im Streben der Physik, die fundamentalen Bausteine der Natur zu beschreiben, stellt die Ent\-wicklung des Standardmodells der Teilchenphysik einen Meilenstein dar. Dieses beschreibt alle bekannten Elementarteilchen sowie deren nicht-gravitative Wechselwirkungen. Neben diesen Erfolgen ist aber auch bekannt, dass es unvollständig ist, da es keine Gravitation oder dunkle Materie beinhaltet. Supersymmetrie ist eine attraktive Erweiterung des Standardmodells, weil sie viele Probelme desselben auf eine elegante Weise löst.\\
Diese Arbeit untersucht eine supersymmetrische Erweiterung des Standardmodells welche zusätzlich $R$-symmetrisch ist. Neben der Berechnung von Wirkungsquerschnitten von stark wechelwirkenden supersymmetrischen Teilchen in der führenden Ordnung werden Einschleifen\-korrekturen zum Produktionswirkungsquerschnitt von Squarks in diesem Model berechnet und phenomenologisch ausgewertet.
\vspace{10em}
\section*{\centering{Abstract}}
In the ambition of understanding the fundamental mechanisms of nature, the Standard Model of particle physics constitutes a cornerstone in physics. It describes all known elementary particles and their non-gravitational interactions. Besides these successes it is known to be incomplete since it does not include gravity or dark matter. Supersymmetry is an attractive extension of the Standard Model as it solves many problems of it elegantly.\\
Within this thesis, a supersymmetric model which also exhibits $R$-symmetry is considered. Besides the calculation of tree-level cross sections for the production of strongly interacting supersymmetric particles, also one-loop corrections to the production cross section of squarks are calculated and phenomenologically analyzed.
\newpage

% Leerseite
\thispagestyle{empty}
\cleardoublepage

\tableofcontents
\newpage

\pagenumbering{arabic}
\input{Introduction}
\newpage
\input{Standardmodel}
\newpage
\input{MSSM}
\newpage
\input{RSymmetry}
\newpage
\input{TreeLevel}
\newpage
\input{Virt_Real}
\newpage
\input{Renormalization}
\newpage
\input{NLO}
\newpage
\input{Summary}
\newpage
% Leerseite
%\thispagestyle{empty}
\cleardoublepage

%\appendix
\input{Appendix}
\newpage
\bibliography{references}
\newpage
\section*{\centering{Acknowledgements}}
First of all I like to thank Dominik Stöckinger who enthused me with theoretical particle physics and provided me with this interesting topic. In addition, he was always available for helping discussions whenever I was in need of them.\\
Special thanks go to Philip Dießner and Wojciech Kotlarski, who contributed significantly to results within this thesis. In addition I like to thank them, Markus Bach, Thomas Kwasnitza, Josua Unger and Tom Steudtner for many valuable discussions and support regarding physical questions as well as programming issues and a pleasant atmosphere in the office. Furthermore, I like to give them thanks for proofreading this thesis.\\
I am particularly grateful for Rebecca Schmidt's support within the last month which enabled me to focus on my thesis. In addition I like to thank her for proofreading.\\
Last but not least I would like to thank my family. Without them my physics studies would by no means have been possible in the way they were.
\newpage
% Erklärung
\clearpage
\thispagestyle{empty}
\minisec{Erklärung}\vspace*{1.5em}
Hiermit erkläre ich, dass ich diese Arbeit im Rahmen der Betreuung am Institut
für Kern- und Teilchenphysik ohne unzulässige Hilfe Dritter verfasst und alle Quellen als solche gekennzeichnet habe.
\vspace*{45em}
\par
Sebastian Liebschner\\
Dresden, Juli 2016

\end{document}